---
title: "MB_SexDiff_GutCheck"
author: "Rachel Moss"
date: "1/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load required packages
```{r LoadPackage, message=FALSE, warning=FALSE}
## Data: /home/ljmills/public/GSE85212_RAW
## srun --nodes=1 --ntasks-per-node=4 --mem=40g  --time=3:00:00  -p ccgg --pty bash -i
library(minfi)
library(minfiData)
library(readr)
# # library(affy)
# library(GEOquery)
# # library(readr)
# # library(EnsDb.Hsapiens.v79)
library(dplyr)
#library(tidyr)
library(tibble)
# library(ggplot2)
# # library(tidyverse)
# # library(ggpubr)
# library(rstatix)
```

Read in filtered GRset - sex given versus predicted sex discordant samples removed, normalization completed (preprocessFunnorm()), and low quality, cross-reactive, snp containing, and sex chromosome probes removed. Additionally read in the combined pheno file.
```{r readdata}
# GRset.flt <- readRDS("./data/GRset.flt.rds")
beta <- read_csv("./data/betaALL121.csv")
beta <- column_to_rownames(beta, var="X1")

pheno <- read_csv("./data/pheno_combined.csv")


```
Match the order of the beta data to that of the pheno file and then run dmpFinder to find differentially methylated probes between SHH and Group 4 
```{r diffmeth}
# pheno_SHHGrp4 <- filter(pheno, pheno$Subgroup=="SHH" | pheno$Subgroup=="Group4")
# beta_SHHGrp4 <- beta[,pheno_SHHGrp4$geo_accession_Meth]
# beta_SHHGrp4_matched <- beta_SHHGrp4[,match(pheno_SHHGrp4$geo_accession_Meth,colnames(beta_SHHGrp4))]
# 
# write.csv(beta_SHHGrp4_matched, "./data/beta_SHHGrp4.csv")
# write.csv(pheno_SHHGrp4, "./data/pheno_SHHGrp4.csv")

beta_SHHGrp4_matched <- read_csv("./data/beta_SHHGrp4.csv")
pheno_SHHGrp4 <- read_csv("./data/pheno_SHHGrp4.csv")

dmp_SHHGrp4 <- dmpFinder(as.matrix(beta_SHHGrp4_matched), pheno = pheno_SHHGrp4$Subgroup, type = "categorical")
head(dmp_SHHGrp4)
write_csv(dmp_SHHGrp4, "./tables"

```

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
