---
title: "MB Sex Differences in DNA Methylation in subgroups - compare sex-DMPs in SHH subtypes"
author: "Rachel Moss"
date: "Mar 10, 2023"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: inline
---

# Introduction

I will use UpSetR package to compare the Cavalli SHH subtype sex-DMPs and the overall SHH sex-DMPs in an upset plot. 

# Setup

## Load packages

```{r packages, message=FALSE, results=FALSE, warning=FALSE}
library(UpSetR)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(xlsx)
```

### Specify files and paths

These DMP files for each subtype for Cavalli SHH are stored on a shared [google drive](https://drive.google.com/drive/folders/1MrmsVwDTWaDP9Es2Lrcb_AihcxN2Fd3x?usp=share_link).

```{r inputs}
# Set the directory where the processed data will be kept
cavDMPDir <- "/Users/moss0134/Google Drive/Research/SexDiffMedullo/Analysis_RMM_new/Results/Cavalli/"
# newDMPDir <- "/Users/moss0134/Google Drive/Research/SexDiffMedullo/Analysis_RMM_new/Results/Newcastle"
# brainDMP <- "/Users/moss0134/Google Drive/Research/SexDiffMedullo/Healthy Brain /GSE15745/GSE15745.top.table.CRBLM_M-F_withCHR.tsv"
# brain2DMP <- "/Users/moss0134/Google Drive/Research/SexDiffMedullo/Analysis_RMM_new/Results/tables/CBL_Ctrl__GSE134379_MethFullResults.csv"
# cavDEG <- "/Volumes/GoogleDrive/.shortcut-targets-by-id/1Ltbi_PE2R7tMPSb5iRZtW_rWMcXE84zl/Medulloblastoma\ Sex\ Differences\ /Natali\ -\ Archive/Natali\ -\ Gene\ Expression\ Top\ Tables"

# File paths to the DMP files for each subgroup
cavSHH_path <- file.path(cavDMPDir, "Cavalli_SHH_ageacatadj_dmp_MethFullResults.csv")
cavSHH_alpha_path <- file.path(cavDMPDir, "/SHH_subtype/Cavalli_SHH_SHH_alpha_ageacatadj_dmp_MethFullResults.csv")
cavSHH_beta_path <- file.path(cavDMPDir, "/SHH_subtype/Cavalli_SHH_SHH_beta_ageacatadj_dmp_MethFullResults.csv")
cavSHH_gamma_path <- file.path(cavDMPDir, "/SHH_subtype/Cavalli_SHH_SHH_gamma_ageacatadj_dmp_MethFullResults.csv")
cavSHH_delta_path <- file.path(cavDMPDir, "/SHH_subtype/Cavalli_SHH_SHH_delta_ageacatadj_dmp_MethFullResults.csv")

#File paths to all the top table files for differential expression done by Natali - /Volumes/GoogleDrive/.shortcut-targets-by-id/1Ltbi_PE2R7tMPSb5iRZtW_rWMcXE84zl/Medulloblastoma\ Sex\ Differences\ /Natali\ -\ Archive/Natali\ -\ Gene\ Expression\ Top\ Tables
# DEG_ALL <- file.path(cavDEG, "diffExp_M-F_ALLnoSexChr.xlsx")
# DEG_SHH <- file.path(cavDEG, "diffExp_M-F_SHHnoSexChr.xlsx")
# DEG_WNT <- file.path(cavDEG, "diffExp_M-F_WNTnoSexChr.xlsx")
# DEG_G3 <- file.path(cavDEG, "diffExp_M-F_G3noSexChr.xlsx")
# DEG_G4 <- file.path(cavDEG, "diffExp_M-F_G4noSexChr.xlsx")



# # Set the directory where the counts files are found
# counts_dir <- here::here("analyses", "mutation_counts")
# 
# # File paths to the counts files for LGAT and medulloblastoma
# lgat_path <- file.path(counts_dir, "LGAT_gene-mutations.tsv")
# mb_path <- file.path(counts_dir, "Medulloblastoma_gene-mutations.tsv")
# 
# # Reference directory & FLAGS (frequently mutated genes)
# ref_dir <- here::here("data", "reference")
# flags_path <- file.path(ref_dir, "FLAGS.txt")
```

Outputs
```{r outputs}
# set the directory to save output plots
results_dir <- "/Users/moss0134/Google Drive/Research/SexDiffMedullo/Analysis_RMM_new/Results/compareDMPs"

# create the results directory if it does not exist
if (!dir.exists(results_dir)){
  dir.create(results_dir, recursive = TRUE)
}
```

### Download Data
Data previously downloaded

```{r downloads}

```

### Define functions
No functions right now

```{r functions}

```

### Reading data

We'll first read in the DMP files data for Cavalli, Newcastle, and healthy brain as tables. For the upset plot, I will have to use the gene names rather than the 450K probeID because different arrays were used. So I clean up the gene names in the DMP files that have duplicates in entries separated by ; because of multiple transcripts. I will also remove any duplicate genes due to probes associated to the same gene. Finally, I'll filter the healthy brain data to those DMPs with adjusted q-val < 0.05 and also removed duplicate gene names.

```{r read_data}
cavSHH <- read_csv(cavSHH_path)
cavSHHa <- read_csv(cavSHH_alpha_path)
cavSHHb <- read_csv(cavSHH_beta_path)
cavSHHd <- read_csv(cavSHH_delta_path)
cavSHHg <- read_csv(cavSHH_gamma_path)

cavSHH <- cavSHH %>% separate_rows(UCSC_RefGene_Name) %>% distinct()
cavSHHa <- cavSHHa %>% separate_rows(UCSC_RefGene_Name) %>% distinct()
cavSHHb <- cavSHHb %>% separate_rows(UCSC_RefGene_Name) %>% distinct()
cavSHHd <- cavSHHd %>% separate_rows(UCSC_RefGene_Name) %>% distinct()
cavSHHg <- cavSHHg %>% separate_rows(UCSC_RefGene_Name) %>% distinct()

cSHH_L <- unique(cavSHH$UCSC_RefGene_Name)
cSHHa_L <- unique(cavSHHa$UCSC_RefGene_Name)
cSHHb_L <- unique(cavSHHb$UCSC_RefGene_Name)
cSHHd_L <- unique(cavSHHd$UCSC_RefGene_Name)
cSHHg_L <- unique(cavSHHg$UCSC_RefGene_Name)

cSHH_Lt <- data.frame("symbol"=cSHH_L,"Cav SHH"=1)
cSHHa_Lt <- data.frame("symbol"=cSHHa_L,"Cav SHH_alpha"=1)
cSHHb_Lt <- data.frame("symbol"=cSHHb_L,"Cav SHH_beta"=1)
cSHHd_Lt <- data.frame("symbol"=cSHHd_L,"Cav SHH_delta"=1)
cSHHg_Lt <- data.frame("symbol"=cSHHg_L, "Cav SHH_gamma"=1)

listI <- list(cSHH_Lt,cSHHa_Lt,cSHHb_Lt,cSHHd_Lt,cSHHg_Lt) %>% purrr::reduce(full_join, by = "symbol")

# remove NA gene found in the brain data
listI <- listI[-6,]
# replace all NAs with 0 in the table
listI[is.na(listI)] <- 0



ListO <- list(cSHH_Lt,cSHHa_Lt,cSHHb_Lt,cSHHd_Lt,cSHHg_Lt) %>% purrr::reduce(full_join, by = "symbol")
# replace all NAs with 0 in the table
ListO[is.na(ListO)] <- 0

cavSHH_p <- cavSHH$probeID
cavSHHa_p <- cavSHHa$probeID
cavSHHb_p <- cavSHHb$probeID
cavSHHd_p <- cavSHHd$probeID
cavSHHg_p <- cavSHHg$probeID

cSHH_pt <-  data.frame("probe"=cavSHH_p, "Cav SHH" = 1)
cSHHa_pt <- data.frame("probe"=cavSHHa_p, "Cav SHH_alpha" = 1)
cSHHb_pt <- data.frame("probe"=cavSHHb_p, "Cav SHH_beta" = 1)
cSHHd_pt <- data.frame("probe"=cavSHHd_p, "Cav SHH_delta" = 1)
cSHHg_pt <- data.frame("probe"=cavSHHg_p, "Cav SHH_gamma" = 1)

listI2 <- list(cSHH_pt,cSHHa_pt,cSHHb_pt,cSHHd_pt,cSHHg_pt) %>% purrr::reduce(full_join, by = "probe")

# replace all NAs with 0 in the table
listI2[is.na(listI2)] <- 0

# cDEG_ALL <- read.xlsx(DEG_ALL,1)
# cDEG_ALL <- filter(cDEG_ALL, adj.P.Val<0.05)
# cDEG_SHH <- read.xlsx(DEG_SHH,1)
# cDEG_SHH <- filter(cDEG_SHH, adj.P.Val<0.05)
# cDEG_WNT <- read.xlsx(DEG_WNT,1)
# cDEG_WNT <- filter(cDEG_WNT, adj.P.Val<0.05)
# cDEG_G3 <- read.xlsx(DEG_G3,1)
# cDEG_G3 <- filter(cDEG_G3, adj.P.Val<0.05)
# cDEG_G4 <- read.xlsx(DEG_G4,1)
# cDEG_G4 <- filter(cDEG_G4, adj.P.Val<0.05)
# 
# cALL_DEG <- data.frame("symbol"=cDEG_ALL$symbol, "DEG All" = 1)
# cSHH_DEG <- data.frame("symbol"=cDEG_SHH$symbol,"DEG SHH"=1)
# cWNT_DEG <- data.frame("symbol"=cDEG_WNT$symbol,"DEG WNT"=1)
# cG3_DEG <- data.frame("symbol"=cDEG_G3$symbol,"DEG Group3"=1)
# cG4_DEG <- data.frame("symbol"=cDEG_G4$symbol,"DEG Group4"=1)
# 
# listDEG <- list(cALL_DEG,cSHH_DEG,cWNT_DEG,cG3_DEG,cG4_DEG) %>% purrr::reduce(full_join, by = "symbol")

```

## Create upset plot
Now I create my upset plot using these lists.
```{r upset, out.width="100%"}
listInput1 <- list("Cavalli SHH" = cSHH_L, "Cavalli SHH_alpha" = cSHHa_L, "Cavalli SHH_beta" = cSHHb_L, "Cavalli SHH_gamma" = cSHHg_L, "Cavalli SHH_delta" = cSHHd_L)

listInput2 <- list("Cavalli SHH" = cSHH_L, "Cavalli SHH_alpha" = cSHHa_L, "Cavalli SHH_beta" = cSHHb_L, "Cavalli SHH_gamma" = cSHHg_L, "Cavalli SHH_delta" = cSHHd_L)

p1 <- upset(fromList(listInput1), sets=c("Cavalli SHH_delta","Cavalli SHH_gamma","Cavalli SHH_beta","Cavalli SHH_alpha","Cavalli SHH"), order.by="degree", keep.order=TRUE, mb.ratio = c(0.55, 0.45),text.scale = c(1.3, 1.3, 1, 1, 2, 1.3))
p1

p2 <- upset(fromList(listInput2), sets=c("Cavalli SHH_delta","Cavalli SHH_gamma","Cavalli SHH_beta","Cavalli SHH_alpha","Cavalli SHH"), order.by="degree", keep.order=TRUE)
p2

listInput1a <- list("Cavalli SHH" = cSHH_L, "Cavalli SHH_alpha" = cSHHa_L, "Cavalli SHH_beta" = cSHHb_L, "Cavalli SHH_gamma" = cSHHg_L, "Cavalli SHH_delta" = cSHHd_L)

p1a <- upset(fromList(listInput1a), sets=c("Cavalli SHH_delta","Cavalli SHH_gamma","Cavalli SHH_beta","Cavalli SHH_alpha","Cavalli SHH"), order.by="degree", keep.order=TRUE, mb.ratio = c(0.55, 0.45),text.scale = c(1.3, 1.3, 1, 1, 2, 1.3))
p1a

```

Since the second set of brain data was 450K, this time I will match by probeID.
```{r upset450k, out.width="100%"}
# listInput3 <- list("Healthy Brain" = brain_p, "Cavalli SHH" = cavSHH_p, "Cavalli WNT" = cavWNT_p, "Cavalli Group3" = cavG3_p, "Cavalli Group4" = cavG4_p, "Newcastle SHH" = newSHH_p, "Newcastle WNT" = newWNT_p, "Newcastle Group3" = newG3_p, "Newcastle Group4" = newG4_p)
# 
# listInput4 <- list("Healthy Brain" = brain_p, "Cavalli ALL" = cavALL_p, "Cavalli SHH" = cavSHH_p, "Cavalli WNT" = cavWNT_p, "Cavalli Group3" = cavG3_p, "Cavalli Group4" = cavG4_p, "Newcastle ALL" = newALL_p, "Newcastle SHH" = newSHH_p, "Newcastle WNT" = newWNT_p, "Newcastle Group3" = newG3_p, "Newcastle Group4" = newG4_p)
# 
# listInput5 <- list("Healthy Brain" = brain_p, "Cavalli SHH" = cavSHH_p, "Cavalli WNT" = cavWNT_p, "Cavalli Group3" = cavG3_p, "Cavalli Group4" = cavG4_p)
# 
# p3 <- upset(fromList(listInput3), sets=c("Newcastle Group4","Cavalli Group4","Newcastle Group3","Cavalli Group3","Newcastle WNT","Cavalli WNT","Newcastle SHH","Cavalli SHH", "Healthy Brain"), order.by="degree", keep.order=TRUE, mb.ratio = c(0.55, 0.45),text.scale = c(1.3, 1.3, 1, 1, 2, 1.3))
# p3
# 
# p4 <- upset(fromList(listInput4), sets=c("Newcastle Group4","Cavalli Group4","Newcastle Group3","Cavalli Group3","Newcastle WNT","Cavalli WNT","Newcastle SHH","Cavalli SHH", "Newcastle ALL","Cavalli ALL","Healthy Brain"), order.by="degree", keep.order=TRUE)
# p4
# 
# p5 <- upset(fromList(listInput5), sets=c("Cavalli Group4","Cavalli Group3","Cavalli WNT","Cavalli SHH","Healthy Brain"), order.by="degree", keep.order=TRUE)
# p5


subgroups <- c("SHHa","SHHb","SHHd","SHHg")

# uniqueGenes <- function(sub,dir,dmps,brain) {
#   unique <- filter(dmps, !(probeID %in% brain$probeID))
#   genes <- as.data.frame(unique$UCSC_RefGene_Name)
#   colnames(genes) <- c(paste0(sub,"_tumorUniqueDMP_genes"))
#   write_csv(genes, paste0(dir,"/",sub,"_tumorUnique_DMPgenes.csv"))
# }
# 
# lapply(subgroups, uniqueGenes(results_dir))


for (sub in subgroups){
  unique <- filter(get(paste0("cav",sub)), !(probeID %in% brain2$probeID))
  genes <- as.data.frame(unique$UCSC_RefGene_Name)
  colnames(genes) <- c(paste0("#Cavalli",sub,"_tumorUniqueDMP_genes"))
  write_csv(genes, paste0(results_dir,"/Cavalli_",sub,"_tumorUnique_DMPgenes.csv"))
}
uSHH <- filter(cavSHH, !(probeID %in% brain2$probeID))
uSHH_g <- as.data.frame(uSHH$UCSC_RefGene_Name)
colnames(uSHH_g) <- c("#Cavalli_SHH_tumorUniqueDMP_genes")
write_csv(uSHH_g, paste0(results_dir,"/SHH_tumorunique_DMPgenes.csv"))

```

### Print all intersections

```{r intersections}
data_with_intersection <- listI %>%
  unite(col = "intersection", -c("symbol"), sep = "")
data_with_intersection %>%
  group_by(intersection) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
intersections <- data_with_intersection %>%
  group_by(intersection) %>%
  summarise(list = list(symbol)) %>%
  mutate(list = setNames(list, intersection)) %>%
  pull(list)

get_intersect_members <- function (x, ...){
  require(dplyr)
  require(tibble)
  x <- x[,sapply(x, is.numeric)][,0<=colMeans(x[,sapply(x, is.numeric)],na.rm=T) & colMeans(x[,sapply(x, is.numeric)],na.rm=T)<=1]
  n <- names(x)
  x %>% tibble::rownames_to_column() -> x
  l <- c(...)
  a <- intersect(names(x), l)
  ar <- vector('list',length(n)+1)
  ar[[1]] <- x
  i=2
  for (item in n) {
    if (item %in% a){
      if (class(x[[item]])=='integer'){
        ar[[i]] <- paste(item, '>= 1')
        i <- i + 1
      }
    } else {
      if (class(x[[item]])=='integer'){
        ar[[i]] <- paste(item, '== 0')
        i <- i + 1
      }
    }
  }
  do.call(filter, ar) %>% tibble::column_to_rownames() -> x
  return(x)
}

```
## Compare DEG to DMPs to see if any match

We want to understand if the DMPs we found are actually expressed

```{r compareDEG}
phenoDir <- "/Users/moss0134/Google Drive/Research/SexDiffMedullo/Analysis_RMM_new/data/"
phenoCav <- read_csv(paste0(phenoDir,"/pheno_CavalliCombined_subgrouped.csv"))

```


### Saving plots

```{r save_plots}
# png(file=file.path(results_dir, "UpsetPlot_compareDMPs_healthybrain.png"),
# width = 1200, height =750, units = 'px')
# p1
# dev.off()

# ggsave(
#   file.path(results_dir, "UpsetPlot_compareDMPs_healthybrain.png"),
#   p1,
# )

```


## Session information

```{r session_info}
sessionInfo()
```
