---
title: "MB_SexDiff_glm"
author: "Rachel Moss"
date: "1/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load required packages
```{r LoadPackage, message=FALSE, warning=FALSE}
## Data: /home/ljmills/public/GSE85212_RAW
## srun --nodes=1 --ntasks-per-node=4 --mem=40g  --time=3:00:00  -p ccgg --pty bash -i
library(minfi)
library(minfiData)
library(readr)
library(dplyr)
library(tibble)
library(survival)
library(mlbench)
library(caret)
library(boot)
library(glmnet)
library(doMC)
library(VennDiagram)
library(RColorBrewer)
library(ggvenn)
```

Read in filtered GRset - sex given versus predicted sex discordant samples removed, normalization completed (preprocessFunnorm()), and low quality, cross-reactive, snp containing, and sex chromosome probes removed. Additionally read in the combined pheno file.
```{r readdata}
# GRset.flt <- readRDS("./data/GRset.flt.rds")
# beta <- read_csv("./data/betaALL121.csv")
# beta <- column_to_rownames(beta, var="X1")
# 
# pheno <- read_csv("./data/pheno_combined.csv")


```
Match the order of the beta data to that of the pheno file and then run dmpFinder to find differentially methylated probes between SHH and Group 4 
```{r diffmeth}
beta_SHHGrp4_matched <- read_csv("./data/beta_SHHGrp4.csv")
beta_SHHGrp4_matched <- column_to_rownames(beta_SHHGrp4_matched, var="X1")
pheno_SHHGrp4 <- read_csv("./data/pheno_SHHGrp4.csv")
pheno_SHH <- filter(pheno_SHHGrp4, pheno_SHHGrp4$Subgroup=="SHH")
beta_SHH <- beta_SHHGrp4_matched[,pheno_SHH$geo_accession_Meth]
# Remove samples with no data for overall survival
keep <- !(is.na(pheno_SHH$OS..years.))
table(keep)
pheno_SHH_survflt <- pheno_SHH[keep,]
# Remove samples that were censored before 5 years - ie OS < 5 and Dead==0
keep <- !(pheno_SHH_survflt$OS..years.<5 & pheno_SHH_survflt$Dead==0)
table(keep)
pheno_SHH_testnot <- pheno_SHH_survflt[!keep,]
pheno_SHH_survflt <- pheno_SHH_survflt[keep,]
# Label samples with no value for Dead but OS > 5 years as Dead==0
pheno_SHH_survflt2 <- pheno_SHH_survflt %>%
  select(Study_ID:geo_accession_Meth) %>%
  mutate(
    Dead_new = case_when(
      !is.na(Dead) ~ Dead,
      is.na(Dead)   ~ 0
    )
  )
# 
keep <- !is.na(pheno_SHH_survflt2$geo_accession_Meth)
pheno_SHH_survflt2 <- pheno_SHH_survflt2[keep,]

beta_SHH_matched <- beta_SHH[,pheno_SHH_survflt2$geo_accession_Meth]

beta_SHH_flt_t <- t(beta_SHH_matched)

```

```{r glmnet}
registerDoMC(cores = 4)
glmnetFit <- glmnet(x=beta_SHH_flt_t, y=pheno_SHH_survflt2$Dead_new, family="binomial")
glmnetFit
plot(glmnetFit)

cvfit <- cv.glmnet(x=beta_SHH_flt_t, y=pheno_SHH_survflt2$Dead_new, family="binomial", type.measure="class")
```
Results / Plot - Plot the cvfit model, showing Misclassification error versus log(lambda)
```{r plotcv}
plot(cvfit)
#plot(cvfit, main="alpha 0.9'')

```

We see from this plot that the lambda that minimized misclassification error still has ~32-33% error.

Proceed now to get the probes that contribute to this model.
```{r features}
c <- coef(cvfit, s= "lambda.min")[,1]
coef <- data.frame(dead5 = c, probe = names(c))
GOImin1 <- filter(coef, dead5 != 0)
#write.table(GOImin,file = "predictive_probes_min_Dead.txt", sep = "\t", quote = F, col.names = T, row.names = F)
c <- coef(cvfit, s= "lambda.1se")[,1]
coef <- data.frame(dead5 = c, probe = names(c))
GOImin2 <- filter(coef, dead5 != 0)
#write.table(GOImin,file = "predictive_probes_lse_Dead.txt", sep = "\t", quote = F, col.names = T, row.names = F)

```

There are 20 probes that contribute.

Now I'll run the model again for just females and just males and see if the probes overlap.
```{r modelbysex}
pheno_SHH_survM <- filter(pheno_SHH_survflt2, predictedSex=="Male")
pheno_SHH_survF <- filter(pheno_SHH_survflt2, predictedSex=="Female")
beta_SHH_survM <- beta_SHH_matched[,pheno_SHH_survM$geo_accession_Meth]
beta_SHH_survF <- beta_SHH_matched[,pheno_SHH_survF$geo_accession_Meth]

beta_SHH_M_t <- t(beta_SHH_survM)
beta_SHH_F_t <- t(beta_SHH_survF)

registerDoMC(cores = 4)
cvfitM <- cv.glmnet(x=beta_SHH_M_t, y=pheno_SHH_survM$Dead_new, family="binomial", type.measure="class")
cvfitF <- cv.glmnet(x=beta_SHH_F_t, y=pheno_SHH_survF$Dead_new, family="binomial", type.measure="class")
```

We plot each of these models.
```{r plotsex, message=FALSE, warning=FALSE, fig.show="hold", out.width="50%"}

plot(cvfitM, main="Male Survival > 5yrs")
plot(cvfitF, main="Female Survival > 5 yrs")

```

Pull out predictive probes from Male and Female survival GLM models
```{r predictprobes}
cM1 <- coef(cvfitM, s= "lambda.min")[,1]
coefM1 <- data.frame(dead5 = cM1, probe = names(cM1))
GOIminM1 <- filter(coefM1, dead5 != 0)
#write.table(GOIminM1,file = "predictive_probes_min_MaleDead.txt", sep = "\t", quote = F, col.names = T, row.names = F)
cM2 <- coef(cvfitM, s= "lambda.1se")[,1]
coefM2 <- data.frame(dead5 = cM2, probe = names(cM2))
GOIminM2 <- filter(coefM2, dead5 != 0)
#write.table(GOIminM2,file = "predictive_probes_lse_MaleDead.txt", sep = "\t", quote = F, col.names = T, row.names = F)

cF1 <- coef(cvfitF, s= "lambda.min")[,1]
coefF1 <- data.frame(dead5 = cF1, probe = names(cF1))
GOIminF1 <- filter(coefF1, dead5 != 0)
#write.table(GOIminF1,file = "predictive_probes_min_FemaleDead.txt", sep = "\t", quote = F, col.names = T, row.names = F)
cF2 <- coef(cvfitF, s= "lambda.1se")[,1]
coefF2 <- data.frame(dead5 = cF2, probe = names(cF2))
GOIminF2 <- filter(coefF2, dead5 != 0)
#write.table(GOIminF2,file = "predictive_probes_lse_FemaleDead.txt", sep = "\t", quote = F, col.names = T, row.names = F)

```

Now let's make a venn diagram to see how many of the probes overlap between the models.
```{r venn}
x <-  list(
  All = GOImin1[-1,2],
  Male = GOIminM1[-1,2], 
  Female = GOIminF1[-1,2]
  )
ggvenn(
  x, 
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF"),
  stroke_size = 0.5, set_name_size = 4
  )

# myCol <- brewer.pal(3, "Pastel2")
# venn.diagram(
#         x = list(GOImin1[-1,2], GOIminM1[-1,2], GOIminF1[-1,2]),
#         category.names = c("All" , "Male only" , "Female Only"),
#         #filename = 'Predictive5yrsurvival_venn_diagramm.png',
#         output=FALSE,
#         
#         # Output features
#         # imagetype="png" ,
#         # height = 480 , 
#         # width = 480 , 
#         # resolution = 300,
#         # compression = "lzw",
#         
#         # Circles
#         lwd = 2,
#         lty = 'blank',
#         fill = myCol,
#         
#         # Numbers
#         cex = .6,
#         fontface = "bold",
#         fontfamily = "sans",
#         
#         # Set names
#         cat.cex = 0.6,
#         cat.fontface = "bold",
#         cat.default.pos = "outer",
#         cat.pos = c(-27, 27, 135),
#         cat.dist = c(0.055, 0.055, 0.085),
#         cat.fontfamily = "sans",
#         rotation = 1
# )


```

